### 2014-11-11 tex/related.tex
Added subsection on JPF implementation details

\subsection{Symbolic Exection}
Generalized symbolic execution \cite{GSE03} was the first symbolic execution method to handle linked data structures as inputs. Subsequent improvements to the method \cite{Deng:2006,Deng:2007} reduce path explosion by delaying path splitting over aliasing choices for as long as possible. However, since all of the above methods introduce nondeterminism during field reads and writes, they all worsen the path explosion problem to some extent or another. Since our heap representation does not introduce any nondeterminism beyond the level in standard symbolic execution, it will generally create fewer paths than these methods, sometimes by a wide margin.

@@

\subsection{Heap Representations}
A number of other techniques model references using sets of guarded locations \cite{Xie:2005,Cherem:2007,Dillig:2011}. These sets, sometimes known as value sets or value summaries, allow multiple heaps to be represented simultaneously with a higher degree of precision than afforded by shape analysis. However, some of these previous attempts utilize non-local recursive definitions for objects. The impact of using recursive definitions for objects is two-fold. First, these techniques must use approximations to resolve reference inequalities, resulting in false positives in the results. Second, these techniques are incapable of reasoning about heaps with loops in the data structures due to infinite recursion during dereferencing. This means that either a large class of data structures must be avoided, or a large class of behaviors must be ignored. In our heap representation, object descriptions are entirely local, thus avoiding these pitfalls.

>>>>

\subsection{Symbolic Exection}
A number of adaptations to allow symbolic execution to cope with heap data structures have been proposed, beginning with Generalized symbolic execution \cite{GSE03}. Subsequent improvements to the method \cite{Deng:2006,Deng:2007} reduce path explosion by delaying path splitting over aliasing choices for as long as possible. However, since all of the above methods introduce nondeterminism during field reads and writes, they all worsen the path explosion problem to some extent or another. Since our heap representation does not introduce any nondeterminism beyond the level in standard symbolic execution, it will generally create fewer paths than these methods, sometimes by a wide margin.

@@

\subsection{Heap Representations}
A number of other techniques model references using sets of guarded locations \cite{Xie:2005,Cherem:2007,Dillig:2011}. These sets, sometimes known as value sets or value summaries, allow multiple heaps to be represented simultaneously with a higher degree of precision than afforded by shape analysis. However, some of these previous attempts  \cite{Xie:2005,Cherem:2007} utilize non-local recursive definitions for objects. The impact of using recursive definitions for objects is two-fold. First, these techniques must use approximations to resolve reference inequalities, resulting in false positives in the results. Second, these techniques are incapable of reasoning about heaps with loops in the data structures due to infinite recursion during dereferencing. This means that either a large class of data structures must be avoided, or a large class of behaviors must be ignored. In our heap representation, object descriptions are entirely local, thus avoiding these pitfalls. The other work \cite{Dillig:2011}, does not handle structures with loops despite having a local object representation, and has further drawbacks as previously enumerated.

### 2014-11-11 tex/results.tex
Added subsection on JPF implementation details

\subsection{JPF Implementation}
The Java Pathfinder (JPF) symbolic execution engine contains an implementation of \gsetxt{}, which we extended with our own implementation of \symtxt{}.

\subsection{Bounding}
In order to perform a practical evaluation of programs involving recursive data structures, it is generally necessary to impose some sort of bounding to eliminate the number of lazy initializations to a manageable level. While these bounds may be asserted as invariants within the program in-situ, it is more practical to apply bounding via external means. There are several bounding techniques that may be applied to both \gsetxt{} and \symtxt{}, including execution path length bounding, k-bounding and n-bounding. k-bounding and n-bounding both operate by imposing limits on the size of the symbolic input heap, but differ in how such bounds are defined. n-bounding limits the number of objects created by lazy initialization, while k-bounding limits the length of initialized reference chains. k-bounding has the property of exhaustively exploring the full breadth of a given data structure, while n-bounding tends to explore deeper heap shapes while being somewhat less thorough in breadth. In addition, k-bounding and n-bounding may be combined to adjust the desired level of approximation between depth and breadth.

@@

BST
The BinarySearchTree example is somewhat of a mixed bag.

>>>>

\subsection{JPF Implementation}
The Java Pathfinder (JPF) symbolic execution engine contains an implementation of \gsetxt{}, which we extended with our own implementation of \symtxt{}. Our heap representation lends itself to a number of simple performance optimizations, which have been included in the implementation. First, solutions to previous constraint problems are cached and re-used, to avoid sending a large number of queries to the SMT solver. Second, because only small portions of the path condition change from one state to the next, incremental solving is enabled to provide a substantial performance increase. Finally, by examining the location sets involved in a reference compare operation, a large number of reference compare instructions may be dispatched without consulting the solver to begin with.

For practical purposes, the implementation also includes a number of features not shown in the simple language described in section 2. For example, integer operations are required in order to reason about ordered data structures such as binary search trees and red/black trees. The description of these operations is outside the scope of this paper.

\subsection{Bounding}
In order to perform a practical evaluation of programs involving recursive data structures, it is generally necessary to impose some sort of bounding to reduce the number of lazy initializations to a manageable level. While these bounds may be asserted as invariants within the program in-situ, it is more practical to apply bounding via external means. There are several bounding techniques that may be applied to both \gsetxt{} and \symtxt{}, including execution path length bounding, k-bounding and n-bounding. k-bounding and n-bounding both operate by imposing limits on the size of the symbolic input heap, but differ in how such bounds are defined. n-bounding limits the number of objects created by lazy initialization, while k-bounding limits the length of initialized reference chains. k-bounding has the property of exhaustively exploring the full breadth of a given data structure, while n-bounding tends to explore deeper heap shapes while being somewhat less thorough in breadth. In addition, k-bounding and n-bounding may be combined to adjust the desired level of approximation between depth and breadth.

@@

The BinarySearchTree example is somewhat of a mixed bag.
